---
title: "Bayesian regression on T-statistics"
author: "Quentin D. Read"
date: "September 8, 2016"
output: 
  pdf_document:
    toc: TRUE
---

# Introduction

Here is a simple Bayesian regression run on the T-statistics data. It is probably superior to the previous regressions for two reasons. One is that it is Bayesian and can give us a better idea of the uncertainty around our estimates of the regression parameters. The other is that it is probably better to combine any traits that we have into a single model with multivariate response data, so that we can account for the effect of the different environmental variables on the covariance between the traits. There are only two traits with enough data to analyze for the mammals (hind foot length and body weight), but I think it is still better to combine them into a single model. The second part doesn't necessarily require the parameter estimation to be Bayesian, but both those things are probably good to do.

I used the *brms* library in R, which is nice because you can write the model specification similarly to how you would write it for the *lm()* and *lmer()* functions, and it automatically converts it for you into Stan code. It outputs the Stan code so we could potentially tweak that code manually and rerun it. I didn't write in all the second-order interactions this time because they were rarely informative in the simpler model. I tried out a few things with model selection, but I ended up just fitting pretty simple models with 4 predictors for each T-statistic. If anyone thinks it would be a good idea to add the interactions, or do more formal model selection, or has any other suggestions for this analysis, please let me know.

# Model specification and fitting

Below, I load all the data and combine it into a single data frame, then split the data frame up into three, one for each T-statistic. I transformed the predictors into standardized variables and got rid of rows with any missing values.

```{r, tidy = TRUE, warning = FALSE, message = FALSE, results='hide'}
# Load data

library(dplyr)
library(lubridate)
library(cati)
source('~/GitHub/NEON/code/analysis/tstat2longform.r')

load('~/GitHub/NEON/allorganismal_latest.r')
load('~/GitHub/NEON/allsiteplot_latest.r')
load('C:/Users/Q/Dropbox/neon/code/tstats_mam_plots15_iucnpool.r')
load('C:/Users/Q/Dropbox/neon/code/mammalPDbyplotobject.r')

spatialstat <- read.csv('C:/Users/Q/Dropbox/neon/data/external_datasets/NEON_spatial_stats_highres.csv')
neonplotdata <- cbind(neonplotdata, spatialstat) %>% rename(ruggedness = tri)

tmiucn <- tstat2longform_ses(tstats_mam_plots15_iucnpool, bysite = FALSE) %>% 
  left_join(filter(neonplotdata, subtype == 'mammalGrid'))

# Mammals have abundance so we use the Chao1 estimator.

chao <- function(x) {
  xcomm <- table(x$taxonID)
  S_obs <- length(xcomm)
  f1 <- sum(xcomm == 1)
  f2 <- sum(xcomm == 2)
  return(data.frame(chao1 = S_obs + (f1 * (f1 - 1)) / (2 * (f2 + 1))))
}

chao1mammal15 <- mam_capture %>% filter(year(date) == 2015) %>%
  group_by(plotID) %>% do(chao(.))

tmiucn <- left_join(tmiucn, chao1mammal15)
tmiucn <- left_join(tmiucn, mammalPD)

# Multiple response variable regression
dat_tipic_all <- tmiucn %>% filter(trait %in% c('weight', 'hindfootLength'), stat == 'T_IP.IC') %>%
  select(trait, t, siteID, plotID, bio1, bio4, cv_bio1, bio12, bio15, cv_bio12, chao1, mpd.obs.z2015, mntd.obs.z2015, ruggedness) %>%
  filter(complete.cases(.)) %>%
  rename(mpd_z = mpd.obs.z2015, mntd_z = mntd.obs.z2015) %>%
  #mutate(siteID = as.numeric(factor(siteID))) %>%
  mutate_each(funs((. - mean(.))/sd(.)), bio1:ruggedness) %>%
  split(.$trait, drop =TRUE)
dat_tipic_all <- left_join(rename(dat_tipic_all[[1]], hindfootLength_t = t)[,-1], rename(dat_tipic_all[[2]], weight_t = t)[,2:4])  

dat_ticir_all <- tmiucn %>% filter(trait %in% c('weight', 'hindfootLength'), stat == 'T_IC.IR') %>%
  select(trait, t, siteID, plotID, bio1, bio4, cv_bio1, bio12, bio15, cv_bio12, chao1, mpd.obs.z2015, mntd.obs.z2015, ruggedness) %>%
  filter(complete.cases(.)) %>%
  rename(mpd_z = mpd.obs.z2015, mntd_z = mntd.obs.z2015) %>%
  #mutate(siteID = as.numeric(factor(siteID))) %>%
  mutate_each(funs((. - mean(.))/sd(.)), bio1:ruggedness) %>%
  split(.$trait, drop =TRUE)
dat_ticir_all <- left_join(rename(dat_ticir_all[[1]], hindfootLength_t = t)[,-1], rename(dat_ticir_all[[2]], weight_t = t)[,2:4])

dat_tpcpr_all <- tmiucn %>% filter(trait %in% c('weight', 'hindfootLength'), stat == 'T_PC.PR') %>%
  select(trait, t, siteID, plotID, bio1, bio4, cv_bio1, bio12, bio15, cv_bio12, chao1, mpd.obs.z2015, mntd.obs.z2015, ruggedness) %>%
  filter(complete.cases(.)) %>%
  rename(mpd_z = mpd.obs.z2015, mntd_z = mntd.obs.z2015) %>%
  #mutate(siteID = as.numeric(factor(siteID))) %>%
  mutate_each(funs((. - mean(.))/sd(.)), bio1:ruggedness) %>%
  split(.$trait, drop =TRUE)
dat_tpcpr_all <- left_join(rename(dat_tpcpr_all[[1]], hindfootLength_t = t)[,-1], rename(dat_tpcpr_all[[2]], weight_t = t)[,2:4])

```

Below, I specify and fit the models. I printed the summary objects as well, which show the parameter estimates and their credible intervals. There are also sampling diagnostics, the effective sample size and Rhat. All parameters have high effective sample size and Rhat of close to 1, which both mean that the chains behaved well.

```{r, tidy = TRUE, eval = FALSE}
library(brms)

brmreduced_tipic <- brm(cbind(weight_t, hindfootLength_t) ~ 0 + trait:(bio1 + chao1 + mpd_z + ruggedness) + (1|siteID), data = dat_tipic_all, chains=4, cores=4, iter = 10000, warmup = 5000)
brmreduced_ticir <- brm(cbind(weight_t, hindfootLength_t) ~ 0 + trait:(bio1 + chao1 + mpd_z + ruggedness) + (1|siteID), data = dat_ticir_all, chains=4, cores=4, iter = 10000, warmup = 5000)
brmreduced_tpcpr <- brm(cbind(weight_t, hindfootLength_t) ~ 0 + trait:(bio1 + chao1 + mpd_z + ruggedness) + (1|siteID), data = dat_tpcpr_all, chains=4, cores=4, iter = 10000, warmup = 5000)
```

```{r, tidy = TRUE, message = FALSE}
library(brms)
load('C:/Users/Q/Dropbox/neon/code/brmreducedobjects.r')

sum_tipic <- summary(brmreduced_tipic)
sum_ticir <- summary(brmreduced_ticir)
sum_tpcpr <- summary(brmreduced_tpcpr)

# Summary for T_IP/IC
sum_tipic

# Summary for T_IC/IR
sum_ticir

# Summary for T_PC_PR
sum_tpcpr

```

\newpage
# Plot of regression coefficients

Below are the regression coefficients for $T_{IP/IC}$ and $T_{IC/IR}$. All the regression coefficients for $T_{PC/PR}$ overlapped zero, so I did not plot them.

```{r, message = FALSE, echo = FALSE, warning = FALSE, fig.height = 7, fig.width = 7}
# Make plots
library(ggplot2)

summ2plotdat <- function(summ) {
  res <- summ$fixed
  params <- strsplit(dimnames(res)[[1]], ':')
  res <- data.frame(predictor = sapply(params, '[', 2),
             trait = sapply(params, '[', 1),
             res)
  dplyr::rename(res, ci_lower = l.95..CI, ci_upper = u.95..CI)
}

pdat <- rbind(transform(summ2plotdat(sum_tipic), stat='Tipic'), transform(summ2plotdat(sum_ticir), stat='Ticir'))
pdat <- transform(pdat, not_zero = (ci_lower<0 & ci_upper<0) | (ci_lower>0 & ci_upper>0))

lab1 <- labeller(stat = c(Tipic = 'Individuals within communities', Ticir = 'Communities within regions'), trait = c(traithindfootLength_t = 'Hind foot length', traitweight_t = 'Body weight'))

library(extrafont)

pcoef <- ggplot(pdat, aes(x = predictor, y = Estimate, ymin = ci_lower, ymax = ci_upper)) +
  geom_hline(yintercept = 0, color = 'blue', linetype = 'dotted') +
  geom_errorbar(width = 0.1) +
  geom_point(aes(color = not_zero, size = not_zero)) +
  facet_grid(stat ~ trait, labeller = lab1) +
  coord_flip() +
  scale_color_manual(values = c('black', 'indianred')) +
  scale_size_manual(values = c(2,3)) +
  scale_x_discrete(labels = c('Temperature', 'Species\nrichness', 'Phylogenetic\ndiversity', 'Topographic\nheterogeneity')) +
  theme_minimal() + theme(legend.position = 'none', panel.border = element_rect(fill = 'transparent'), text = element_text(family = 'Century Gothic'), axis.title.y = element_blank())

pcoef

```

