---
title: "Phylogenetic diversity methods"
author: "Quentin D. Read"
date: "August 25, 2016"
output: pdf_document
---

```{r, message=FALSE, echo=FALSE, warning=FALSE, results='hide'}
# Load in the phylogeny for mammals

library(ape)
library(picante)
tol <- read.tree('C:/Users/Q/Dropbox/neon/data/external_datasets/phylo/TimetreeOfLife2015.nwk')

# Load mammal names
mammalTax <- read.csv('~/GitHub/NEON/species_lists/mammalSpeciesList.csv', stringsAsFactors = FALSE)

# Load NEON data and calculate phylogenetic diversity indices
load('~/GitHub/NEON/allorganismal_latest.r')

# Process data to get communities for each year (same communities as are used for the Chao calculation)

library(dplyr)
library(lubridate)
library(reshape2)

# These are all the capture grids, with possibly multiple bouts per year
# Get rid of blank or "Other" taxonIDs, and two individuals with a possibly typo species code
mam_comms <- mam_capture %>% mutate(gridID = substr(trapCoordinate,1,1), year = year(date)) %>%
  filter(taxonID != '', taxonID != 'OTHE', taxonID != 'ROSP') %>%
  group_by(year, siteID) %>%
  do(comm = dcast(., formula = gridID ~ taxonID)[,-1, drop = FALSE]) # Get rid of date col and use n. rows as n. individuals

mam_comms_byyear <- do(mam_comms, abund = apply(.$comm, 2, sum))

# Make a single matrix for each year.
comm_mat <- function(x) {
  mat <- dcast(x, formula = siteID ~ taxonID)
  row.names(mat) <- mat$siteID
  mat[,-1, drop=FALSE]
}

mam_mats <- mam_capture %>% mutate(year = year(date)) %>%
  filter(taxonID != '', taxonID != 'OTHE', taxonID != 'ROSP') %>%
  group_by(year) %>%
  do(comm = comm_mat(.))

# Get the list of four-letter codes that are in the mammal matrices
allIDs <- Reduce(union, sapply(mam_mats$comm, names))
allIDs <- merge(data.frame(taxonID=allIDs), mammalTax[,c('taxonID','scientificName')], all.x=TRUE)
allIDs <- transform(allIDs, scientificName = sub(" ", "_", scientificName))

# Figure out which of the mammals in mammal community aren't in the tree, and add them to "tol" using phyTools
# It is a lot of unidentified species, and some Spermophilus species.

spp_not_in_tree <- allIDs$scientificName[!allIDs$scientificName %in% tol$tip.label]

# Extract the mammal clade from phylogeny
sum(mammalTax$scientificName %in% tol$tip.label) # Wrong format
sum(with(mammalTax, paste(genus, specificEpithet, sep='_')) %in% tol$tip.label) # 300 are in

not_mammals <- which(!(tol$tip.label %in% allIDs$scientificName) & !grepl('Spermophilus', tol$tip.label))
mammaltol <- drop.tip(tol, tip=not_mammals) 

# Add tips for individuals that are only identified to genus, and add Spermophilus.
library(phytools)
for (i in spp_not_in_tree[1:9]) mammaltol <- add.species.to.genus(tree = mammaltol, species = i, where = 'root')
for (i in spp_not_in_tree[10:13]) mammaltol <- add.species.to.genus(tree = mammaltol, species = i, where = 'random')

# Prune out the other spermophilus entries
mammaltol <- drop.tip(mammaltol, tip=which(!(mammaltol$tip.label %in% allIDs$scientificName)))

# Change the tip labels of the mammal phylogeny to the taxonIDs
mammaltol$tip.label <- as.character(allIDs$taxonID)[match(mammaltol$tip.label, allIDs$scientificName)]


# Calculate pd metrics ----------------------------------------------------

# PD
mc2014 <- mam_mats$comm[[2]]
mc2015 <- mam_mats$comm[[3]]

tree2014 <- prune.sample(mc2014, mammaltol)
tree2015 <- prune.sample(mc2015, mammaltol)

pd2014 <- pd(mc2014, tree2014)
pd2015 <- pd(mc2015, tree2015)

# MPD and MNTD
dist2014 <- cophenetic(tree2014)
dist2015 <- cophenetic(tree2015)

# Compare null models within MPD and within MNTD

mpds <- list()
mntds <- list()
nms <- c("taxa.labels", "richness", "frequency", "sample.pool", "phylogeny.pool", "independentswap", "trialswap")

for (nm in nms) {
  mpds <- within(mpds, assign(nm, ses.mpd(mc2015, dist2015, null.model = nm, abundance.weighted = TRUE, runs = 999)$mpd.obs.z))
  mntds <- within(mntds, assign(nm, ses.mntd(mc2015, dist2015, null.model = nm, abundance.weighted = TRUE, runs = 999)$mntd.obs.z))
}


```


I am not familiar with phylogenetic diversity metrics, so I wanted to know if anyone had any ideas about which methods are the best to use for our questions. I know that mean pairwise distance and mean nearest taxon distance show two different aspects of diversity: mean pairwise distance describes large-scale patterns in the tree more broadly across the phylogeny, and mean nearest taxon distance picks up the finer differences. However there are a number of randomization methods for the null models with which you calculate the standardized effect sizes of MPD and MNTD (which are the opposite of NRI and NTI).

The names of the methods in the *picante* package are:

* taxa.labels
* richness
* frequency
* sample.pool
* phylogeny.pool
* independentswap
* trialswap

Within MPD and MNTD, I plotted these methods pairwise against one another to show which ones differ, and I also plotted the MPD versus the MNTD for each method to see if the relationship between them was similar depending on which method I used. The plots are below, I would appreciate any feedback on what seems to be going on here and what is the best method to use.

```{r, message=FALSE, echo=FALSE, fig.height=7, fig.width=7}
library(GGally)
ggpairs(do.call('cbind', mpds))
```

**Figure 1.** Pairwise plots of SES.MPD (standardized effect size of mean pairwise distance, or -1 * NRI), by randomization method.

```{r, message=FALSE, echo=FALSE, fig.height=7, fig.width=7}
ggpairs(do.call('cbind', mntds))
```

**Figure 2.** Pairwise plots of SES.MNTD (standardized effect size of mean nearest taxon distance, or -1 * NTI), by randomization method.

\newpage
```{r, message=FALSE, echo=FALSE, fig.height=3, fig.width=3}
library(ggplot2)
source('~/qutil.r')
ll <- geom_abline(slope=1, color = 'red', linetype = 'dotted') 
qplot(mpds$taxa.labels, mntds$taxa.labels) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'taxa.labels Randomization')
qplot(mpds$richness, mntds$richness) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'richness Randomization')
qplot(mpds$frequency, mntds$frequency) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'frequency Randomization')
qplot(mpds$sample.pool, mntds$sample.pool) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'sample.pool Randomization')
qplot(mpds$phylogeny.pool, mntds$phylogeny.pool) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'phylogeny.pool Randomization')
qplot(mpds$independentswap, mntds$independentswap) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'independentswap Randomization')
qplot(mpds$trialswap, mntds$trialswap) + ll + coord_equal() +
  qSubtitle('MNTD versus MPD', 'trialswap Randomization')
```

**Figure 3.** MNTD versus MPD for all methods. The y=x line is the red dotted line.